The \dealii{} Library, Version 8.5

### 2017-03-09 8.5/paper.tex
Restructure and reword Section 2.3 LinearOperator

\item The \texttt{LinearOperator} class has been extended by a generic
  ``payload'' mechanism that allows to attach arbitrary additional
  information to a \texttt{LinearOperator}. The mechanism was used to
  implement a \texttt{TrilinosPayload} class to have full support for using
  the suite of {\trilinos} iterative solvers and preconditioners with a
  \texttt{LinearOperator}.

@@

This release of \dealii{} contains a number of large and significant changes
that will be discussed in the following sections. It of course also contains a
vast number of smaller changes and added functionality; the details of these
can be found
\href{https://www.dealii.org/8.5.0/doxygen/deal.II/changes_between_8_4_and_8_5.html}{in the file that lists all changes for this release} (see \cite{changes85})
and that is linked to from the web site of each release as well as the
release announcement.

@@

The \texttt{CellDataStorage} class is an integrated mechanism to safely store user-defined data, such as that required at each computation point, within each cell.
Although the same was previously achieved through the use of a cell \texttt{user\_pointer}, it required users to manage this data themselves.
As this data is now treated as a first-class citizen to \dealii{}, through the abstract \texttt{TransferableQuadraturePointData} we have provided a generic interface to facilitate several low-level operations that were previously tedious or technically challenging to implement.
For instance, the \texttt{parallel::distributed::ContinuousQuadratureDataTransfer} class assists in the transfer of arbitrary data (that is continuous within a cell) stored at quadrature points when performing h-adaptive refinement of \texttt{parallel::distributed::Triangulation}.
Not only does it perform an $\mathcal{L}^2$ projection of the specified user data between quadrature points, but it also ships the data automatically between MPI processes.

@@

By introduction of an arbitrary \texttt{Payload} base class, the
\texttt{LinearOperator} class can be extended to seamlessly exploit the
native features and operations offered by external linear algebra
libraries. We have thus developed a \texttt{TrilinosPayload} class, through
which full support for the suite of {\trilinos} parallel iterative solvers
and preconditioners has been achieved. In the case of the construction of
an \texttt{inverse\_operator}, both \dealii's built in solvers and those
offered by {\trilinos} can be selected. This has been achieved by using the
\texttt{Epetra\_Operator} as the basis for the \texttt{TrilinosPayload},
for which the result of both standard and composite operations involving
forward (\texttt{Apply()}) and inverse (\texttt{ApplyInverse()})
matrix-vector multiplication are collated using lambda functions. It is
envisaged that, in the future, similar can be done for the {\petsc}
iterative solvers.

An additional feature of the \texttt{LinearOperator} suite is the defintion
of a \texttt{schur\_complement} operator, and its associated condensation
and post-processing \texttt{PackagedOperation}s. An operator representing
the Schur complement of a block system can be declared and, though the
delayed evaluation offered by \texttt{PackagedOperation}s, reused on any
number of vector systems.

>>>>

\item the \texttt{LinearOperator} class has been extended by a generic
  ``payload'' mechanism that allows to attach arbitrary additional
  information to a \texttt{LinearOperator}. The mechanism was used to
  implement a \texttt{TrilinosPayload} class to have full support for using
  the suite of {\trilinos} iterative solvers and preconditioners with a
  \texttt{LinearOperator}.

@@

This release of \dealii{} contains a number of large and significant changes
that will be discussed in the following sections. It of course also contains a
vast number of smaller changes and added functionality; the details of these
can be found
\href{https://www.dealii.org/8.5.0/doxygen/deal.II/changes_between_8_4_and_8_5.html}{in
the file that lists all changes for this release}; see \cite{changes85}.
(The file is also linked to from the web site of each release as well as
the release announcement.)

@@

The \texttt{CellDataStorage} class is an integrated mechanism to safely
store user-defined data, such as that required at each computation point,
within each cell. Although the same was previously achieved through the use
of a cell \texttt{user\_pointer}, it required users to manage this data
themselves. This data is now treated as a first-class citizen to \dealii{}.

Through the abstract \texttt{TransferableQuadraturePointData} we have
provided a generic interface to facilitate several low-level operations
that were previously tedious or technically challenging to implement. For
instance, the
\texttt{parallel::distributed::ContinuousQuadratureDataTransfer} class
assists in the transfer of arbitrary data (that is continuous within a
cell) stored at quadrature points when performing h-adaptive refinement of
\texttt{parallel::distributed::Triangulation}. Not only does it perform an
$\mathcal{L}^2$-projection of the specified user data between quadrature
points, but it also ships the data automatically between MPI processes.

@@

The \texttt{LinearOperator} class has been extended by a generic
``payload'' mechanism that allows to attach arbitrary additional
information to a \texttt{LinearOperator}. This was achieved by introducing
a generic \texttt{Payload} base class. The main use case of the new
mechanism is to extend the \texttt{LinearOperator} class to seamlessly
exploit the native features and operations offered by external linear
algebra libraries. We have thus developed a \texttt{TrilinosPayload} class,
that provides full support for the suite of {\trilinos} parallel iterative
solvers and preconditioners.

A particular interesting case is the construction of an
\texttt{inverse\_operator}. Now both, \dealii's built in solvers and those
offered by {\trilinos}, can be selected. This has been achieved by using
the \texttt{Epetra\_Operator} as the basis for the
\texttt{TrilinosPayload}, for which the result of standard and composite
operations involving forward (\texttt{Apply()}) and inverse
(\texttt{ApplyInverse()}) matrix-vector multiplication are collated using
lambda functions.

It is envisaged that, in the future, similar can be done for the {\petsc}
iterative solvers.

An additional feature of the \texttt{LinearOperator} suite is the defintion
of a \texttt{schur\_complement} operator and its associated condensation
and post-processing \texttt{PackagedOperation}s. An operator representing
the Schur complement of a block system can be declared and, though the
delayed evaluation offered by \texttt{PackagedOperation}s, reused on any
number of vector systems.
